---
title: Project 2 Report 
date: "Due 5 pm PST March 5th 2021"
author: "Group 1"
output:
  html_document:
    df_print: paged
    numbered_sections: yes
---

```{r global_options, include=FALSE}
# Setup this markdown file:
knitr::opts_chunk$set(
  fig.pos = 'H',
  warning = FALSE,
  message = FALSE,
  eval = TRUE, 
  echo = TRUE, 
  fig.align='center'
)
# Include all necessary libraries
library(tidyverse)
library(usmap)
library(reshape2)
```

***

Reproducibility:
- The source file  (e.g., RMD file)  is well-documented with reproducibility in mind. For instance, a readme document is included to offer instructions for using these files, codes are well-commented, the files are well-structured, the script/file are tested in various systems, etc.

# Introduction
The effects of the COVID-19 pandemic on mental health have become an increasingly recognized issue in public discourse, yet mental health science perspectives remain relatively overlooked in the construction of research priorities and policy agendas related to recovery from the COVID-19 pandemic. Researchers recognize that the psychological and social effects of the COVID-19 pandemic on mental health are pervasive [(Holmes & O’Connor et al., 2020, p. 547)](https://doi.org/10.1016/S2215-0366(20)30168-1) and that, without the support of government programs and policy, individuals are at risk of bearing a “psychological scar” [(Clemente-Suárez et al, 2020, p. 4)](doi: 10.3389/fpsyg.2020.580225) long after the pandemic ends. Therefore, coordination between governments and global research platforms is necessary to further address how mental health consequences for vulnerable groups can be mitigated under pandemic conditions [(Holmes & O’Connor et al., 2020, p. 547)](https://doi.org/10.1016/S2215-0366(20)30168-1).

The objective of this report is to inform readers about the impacts of COVID-19 on mental health in the United States. Although mental health is a global issue, the decentralized and inconsistent response to COVID-19 in the United States motivates us to explore the United States as a case study for populations in countries implementing dynamic responses to the pandemic; the cyclic nature of lockdowns and control measures implies that the stressors associated with COVID-19 will persist in the United States, so knowledge about COVID-19 mental health consequences will help guide effective government responses and assistance programs.

To investigate the impact of COVID-19 on mental health in the United States, we define our primary question of interest as whether an association exists between the number of COVID-19 cases and deaths and the frequency of indicators of depressive and anxiety disorder. Furthermore, we define our secondary question of interest as, if an association exists between the COVID-19 new case and death count and indicators of depressive and anxiety disorder, whether a specific population in the United States experiences the greatest frequency of symptoms.

We analyze data from the [WHO Coronavirus Dashboard](https://covid19.who.int/) and the Household Pulse Survey [(HPS)](https://www.cdc.gov/nchs/covid19/pulse/mental-health.htm). The WHO Coronavirus Dashboard reflects laboratory-confirmed cases and deaths based upon WHO case definitions, although some departures may exist due to local adaptations. The HPS data set documents the percentage of respondents who report symptoms of anxiety or depression that have been shown to be associated with diagnoses of generalized anxiety disorder or major depressive disorder. The key variables pulled from the WHO Coronavirus Dashboard include New Cases and New Deaths to investigate COVID-19 new case and death count. The key variables pulled from the HPS include Indicator, which we use to identify frequency of symptoms related to depressive or anxiety disorder, and Group, which we use to define populations within the United States. 

Prior analysis of HPS data suggests a disproportionate negative impact of COVID-19 stressors on emotional distress among older persons of color [(Bui et al, 2020, p. 262)](doi:10.1093/geront/gnaa217) and a significant mental health burden stemming from job insecurity among young adults amidst the COVID-19 pandemic [(Ganson et al., 2021, p. 55)](https://doi.org/10.1016/j.jadohealth.2020.10.008). Therefore, we hypothesize that the frequency of indicators of depressive or anxiety disorder increases as the new case and death count associated with COVID-19 grows in the United States. We also expect that, when comparing populations defined by ethnicity, non-White groups will show the greatest frequency of symptoms. When comparing populations defined by age, we expect to find the highest frequency of symptoms related to depressive or anxiety disorder among young adults.


Rubric:
- The introduction clearly states the questions of interest and the impact of results of this data analysis, along with real-world motivation. 
- In addition, the introduction provides a precise and concise description of the problem setup, source of data, key variables, and possible hypotheses. 
- Readers that are familiar with this topic can easily understand the introduction and recognize the importance of this project. 

# Background
Rubric:
- The report contains extensive and selective discussion of the source of data, target population, sampling mechanism, explanation of variables, and  reviews of existing research or known results that are relevant to this topic. 
- The background discussion gives readers, who are familiar with this topic or statistics, sufficient information to place this analysis in the literature, e.g., can find research with similar tasks, can compare the results in this report against other results.

# Descriptive Analysis
Rubric:
- The descriptive analysis provides thorough descriptive statistics for most variables, which include univariate descriptive statistics for all variables (mean, standard deviations, missing values,  quantiles, etc.), multivariate descriptives of key variables, and insightful patterns in the data set. The descriptives are presented in visually appealing and informative formats. 
- Readers that are familiar with this topic or statistics understand the major features in this data set, and gain new insights of this data set that are not possible without the descriptives in this report.

```{r}
# Import WHO COVID Data, CDC Depression and Anxiety Data, and Google Trends
covid <- read.csv(file = "WHO-COVID-19-global-data.csv", fileEncoding="UTF-8-BOM")
indicators <- read.csv(file = "CDC_Anxiety.csv", header=T)
GOOGLE <- read.csv(file = "Google_US_Depression_Anxiety.csv")

# Convert date column to POSIXct
GOOGLE$Week <- as.POSIXct(GOOGLE$Week)
GOOGLE <- subset(GOOGLE, Week >= as.POSIXct("2020-01-01"))

# subset covid data to include only US data 
covid_US = subset(covid, Country == "United States of America") %>% select(Date_reported,New_cases,Cumulative_cases,New_deaths,Cumulative_deaths)
covid_US$Date_reported <- as.POSIXct(covid_US$Date_reported, format = "%Y-%m-%d")

# subset CDC anxiety data to include only the measures for anxiety or depressive disorder (combined) and the national measure 
indicators_US = subset(indicators, Indicator == "Symptoms of Anxiety Disorder or Depressive Disorder" & 
                         Group == "National Estimate" ) %>%     select(Time.Period.Label,Time.Period.Start.Date,Time.Period.End.Date,Value,Low.CI,High.CI,Quartile.Range)

# remove na data from the indicators leaving 24 values for 24 distinct time periods  
indicators_US.narm <- indicators_US %>% filter(!is.na(Value))

# Convert start/end date to standard format.
indicators_US.narm$Time.Period.Start.Date <- as.POSIXct(indicators_US.narm$Time.Period.Start.Date, format = "%m/%d/%Y")
indicators_US.narm$Time.Period.End.Date <- as.POSIXct(indicators_US.narm$Time.Period.End.Date, format = "%m/%d/%Y")
```

```{r}
# Present Depression or Anxiety Information by Race
indicators.race <- subset(indicators, Indicator == "Symptoms of Anxiety Disorder or Depressive Disorder" & 
                         Group == "By Race/Hispanic ethnicity", c(Subgroup, Value)) %>% filter(!is.na(Value))
indicators.race$Subgroup <- as.character(indicators.race$Subgroup)
indicators.race.means <- aggregate(indicators.race, list(indicators.race$Subgroup), mean)
ggplot(indicators.race.means, aes(x=Value, y=Group.1, )) + 
  geom_bar(aes(fill=Group.1), stat = "identity", position = "dodge") +
  theme(legend.position = "none") +
  labs(x = "Average % of Survey Responses Indicating Anxiety or Depression", y = "")

# Present Depression or Anxiety Information by Sex
indicators.sex <- subset(indicators, Indicator == "Symptoms of Anxiety Disorder or Depressive Disorder" & 
                         Group == "By Gender",c(Subgroup, Value))  %>% filter(!is.na(Value))
indicators.sex.means <- aggregate(indicators.sex, list(indicators.sex$Subgroup), mean)
ggplot(indicators.sex.means, aes(x=Value, y=Group.1)) + 
  geom_bar(aes(fill=Group.1), stat = "identity", position = "dodge") +
  theme(legend.position = "none") +
  labs(x = "Average % of Survey Responses Indicating Anxiety or Depression", y = "")

# Present Depression or Anxiety Information by State
indicators.state <- subset(indicators, Indicator == "Symptoms of Anxiety Disorder or Depressive Disorder" & 
                         Group == "By State", c(Subgroup,Value))  %>% filter(!is.na(Value))
indicators.state.means <- aggregate(indicators.state, list(indicators.state$Subgroup), mean) %>% select(Group.1,Value)
indicators.state.means$Group.1 <- usmap::fips(indicators.state.means$Group.1)
names(indicators.state.means) <- c('fips','Value')
usmap::plot_usmap(data=indicators.state.means, values="Value") +
  theme(legend.position = 'right') +
  labs(title = "Depression or Anxiety Symptoms", subtitle = "Average 2020 Household Pulse Survey") +
  scale_fill_continuous(low = "yellow", high = "blue", name = "Anxiety or Depression Percentage")
```


```{r}
# Fill in dates of indicators and GOOGLE
HPS <- list()
for (i in 1:(dim(indicators_US.narm)[1]-1)) {
  fill.dates <- seq(indicators_US.narm[i,'Time.Period.Start.Date'], indicators_US.narm[i+1,'Time.Period.Start.Date'] - 1*60*60*24, by="days")
  fill.values <- c(rep(1,length(fill.dates)))*indicators_US.narm[i,'Value']
  fill.dates.df <- data.frame(fill.dates,fill.values)
  HPS[[i]] <- fill.dates.df
}
HPS.full <- bind_rows(HPS)

Trends.Anxiety <- list()
Trends.Depression <- list()
for (i in 1:(dim(GOOGLE)[1]-1)) {
  fill.dates <- seq(GOOGLE[i,'Week'], GOOGLE[i+1,'Week'] - 1*60*60*24, by="days")
  fill.anxiety.values <- c(rep(1,length(fill.dates)))*GOOGLE[i,'Anxiety...United.States.']
  fill.depression.values <- c(rep(1,length(fill.dates)))*GOOGLE[i,'Depression...United.States.']
  fill.dates.df <- data.frame(fill.dates,fill.anxiety.values)
  Trends.Anxiety[[i]] <- fill.dates.df
  fill.dates.df <- data.frame(fill.dates,fill.depression.values)
  Trends.Depression[[i]] <- fill.dates.df
}
Trends.Anxiety.Full <- bind_rows(Trends.Anxiety)
Trends.Depression.Full <- bind_rows(Trends.Depression)

# Combine into one dataframe
all.covid <- covid_US %>% subset(Date_reported >= as.POSIXct("2020-04-23") & Date_reported < as.POSIXct("2021-01-01"),c(Date_reported,New_cases,New_deaths))
all.Anxiety <- Trends.Anxiety.Full %>% subset(fill.dates >= as.POSIXct("2020-04-23") & fill.dates < as.POSIXct("2021-01-01"))
all.Depression <- Trends.Depression.Full %>% subset(fill.dates >= as.POSIXct("2020-04-23") & fill.dates < as.POSIXct("2021-01-01"))
all.HPS <- HPS.full %>% subset(fill.dates >= as.POSIXct("2020-04-23") & fill.dates < as.POSIXct("2021-01-01"))
all.df <- data.frame(all.covid,all.Anxiety$fill.anxiety.values,all.Depression$fill.depression.values,all.HPS$fill.values)
names(all.df) <- c("Date","New Cases","New Deaths","Anxiety Google Trends","Depression Google Trends","Household Pulse Survey")
```

```{r}
# Show daily deaths, cases, anxiety|depression, and google trends overtime
all.df.m <- melt(all.df, "Date")
# New_cases,New_deaths,all.Anxiety.fill.anxiety.values,all.Depression.fill.depression.values,all.HPS.fill.values
ggplot(all.df.m, aes(Date,value,colour = variable)) + 
  geom_point() + 
  geom_point() + 
  geom_point() + 
  geom_point() + 
  geom_point() + 
  facet_wrap(~ variable, ncol=1,scales = "free_y") +
  labs(title = "Depression or Anxiety Symptoms", subtitle = "Average 2020 Household Pulse Survey")
```

# Predictive/Inferential Analysis
Rubric:
- The report clearly states the statistical model, method, or procedure in use, which includes the introduction of notation, interpretation of the parameters, statements of assumptions, etc. There are no major errors in the model and implementation of the model. 
- The report provides sound justification of the chosen model and the associated assumptions in the context of the data set. Analysis results are well-intepreted with appealing visualization that answers the stated questions of interest . 
- Readers that are familiar with this topic or statistics can understand the proposed model/method/procedure and the retionale behind it. Most concerns or disagreements have been addressed in the report. 
- Furthermore, based on the report, readers who have sufficient knowledge in statistics can fully reproduce the analysis on their own.

```{r setup, include=FALSE}
#setting the working directory 
setwd("C:/Users/leakelly/Desktop/STA 141 A")
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, fig.align='center')

#loading the covid data 
covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
covid

#loading the CDC anxiety data
indicators = read.csv("Indicators_of_Anxiety_or_Depression_Based_on_Reported_Frequency_of_Symptoms_During_Last_7_Days.csv", header=TRUE)
indicators

#subsetting the covid data to include only the data from the US
covid_US = subset(covid, Country == "United States of America")
covid_US

#subsetting the CDC anxiety data to include only the measures for anxiety or depressive disorder (combined) and the national measure
indicators_US = subset(indicators, Indicator == "Symptoms of Anxiety Disorder or Depressive Disorder" & 
                         Group == "National Estimate" )
indicators_US

#removing the na data from the indicators dataset, this leaves 24 values for 24 distinct time periods  
indicators_US.narm <- indicators_US %>% filter(!is.na(Value))
```

```{r}
#plot looking at the cumulative deaths over the dates reported
qplot(Date_reported, Cumulative_deaths, data = covid_US)

#plot looking at the cases of covid over the dates reported 
qplot(Date_reported, New_cases, data = covid_US)

#plot looking at the CDC anxiety data over the time reported 
qplot(Time.Period, Value, data = indicators_US.narm)
```

<span style='color:blue'>
In order to create the linear model first the dates needed to be standardized. The CDC Depression and Anxiety data has 24 datapoints for distinct time periods. The WHO COVID-19 data has 427 datapoints for distinct days. The days in the WHO COVID-19 that math ot the time periods in the CDC Depression and Anxiety data were then compiled and averages were taken for new cases and new deaths. Doing this contrils for the time series variable in the data. 
</span>

```{r}
dates1 = covid_US[112:124,]
dates2 = covid_US[126:131,]
dates3 = covid_US[133:138,]
dates4 = covid_US[140:145,]
dates5 = covid_US[147:152,]
dates6 = covid_US[154:159,]
dates7 = covid_US[161:166,]
dates8 = covid_US[168:173,]
dates9 = covid_US[175:180,]
dates10 = covid_US[182:187,]
dates11 = covid_US[189:194,]
dates12 = covid_US[196:201,]
dates13 = covid_US[230:242,]
dates14 = covid_US[244:256,]
dates15 = covid_US[258:270,]
dates16 = covid_US[272:284,]
dates17 = covid_US[286:298,]
dates18 = covid_US[300:312,]
dates19 = covid_US[314:326,]
dates20 = covid_US[328:340,]
dates21 = covid_US[342:354,]
dates22 = covid_US[370:382,]
dates23 = covid_US[384:396,]
dates24 = covid_US[398:410,]

dates1_cases = mean(dates1$New_cases)
dates1_deaths = mean(dates1$New_deaths)

dates2_cases = mean(dates2$New_cases)
dates2_deaths = mean(dates2$New_deaths)

dates3_cases = mean(dates3$New_cases)
dates3_deaths = mean(dates3$New_deaths)

dates4_cases = mean(dates4$New_cases)
dates4_deaths = mean(dates4$New_deaths)

dates5_cases = mean(dates5$New_cases)
dates5_deaths = mean(dates5$New_deaths)

dates6_cases = mean(dates6$New_cases)
dates6_deaths = mean(dates6$New_deaths)

dates7_cases = mean(dates7$New_cases)
dates7_deaths = mean(dates7$New_deaths)

dates8_cases = mean(dates8$New_cases)
dates8_deaths = mean(dates8$New_deaths)

dates9_cases = mean(dates9$New_cases)
dates9_deaths = mean(dates9$New_deaths)

dates10_cases = mean(dates10$New_cases)
dates10_deaths = mean(dates10$New_deaths)

dates11_cases = mean(dates11$New_cases)
dates11_deaths = mean(dates11$New_deaths)

dates12_cases = mean(dates12$New_cases)
dates12_deaths = mean(dates12$New_deaths)

dates13_cases = mean(dates13$New_cases)
dates13_deaths = mean(dates13$New_deaths)

dates14_cases = mean(dates14$New_cases)
dates14_deaths = mean(dates14$New_deaths)

dates15_cases = mean(dates15$New_cases)
dates15_deaths = mean(dates15$New_deaths)

dates16_cases = mean(dates16$New_cases)
dates16_deaths = mean(dates16$New_deaths)

dates17_cases = mean(dates17$New_cases)
dates17_deaths = mean(dates17$New_deaths)

dates18_cases = mean(dates18$New_cases)
dates18_deaths = mean(dates18$New_deaths)

dates19_cases = mean(dates19$New_cases)
dates19_deaths = mean(dates19$New_deaths)

dates20_cases = mean(dates20$New_cases)
dates20_deaths = mean(dates20$New_deaths)

dates21_cases = mean(dates21$New_cases)
dates21_deaths = mean(dates21$New_deaths)

dates22_cases = mean(dates22$New_cases)
dates22_deaths = mean(dates22$New_deaths)

dates23_cases = mean(dates23$New_cases)
dates23_deaths = mean(dates23$New_deaths)

dates24_cases = mean(dates24$New_cases)
dates24_deaths = mean(dates24$New_deaths)

dates = c("dates1", "dates2", "dates3", "dates4", "dates5", "dates 6", "dates7", "dates8", "dates9", "dates10", "dates11", "dates12", "dates13", "dates14", "dates15", "dates16", "dates17", "dates18", "dates19", "dates20", "dates21", "dates22", "dates23", "dates24")

cases = c(dates1_cases, dates2_cases, dates3_cases, dates4_cases, dates5_cases, dates6_cases, dates7_cases, dates8_cases, dates9_cases, dates10_cases, dates11_cases, dates12_cases, dates13_cases, dates14_cases, dates15_cases, dates16_cases, dates17_cases, dates18_cases, dates19_cases, dates20_cases, dates21_cases, dates22_cases, dates23_cases, dates24_cases)

deaths = c(dates1_deaths, dates2_deaths, dates3_deaths, dates4_deaths, dates5_deaths, dates6_deaths, dates7_deaths, dates8_deaths, dates9_deaths, dates10_deaths, dates11_deaths, dates12_deaths, dates13_deaths, dates14_deaths, dates15_deaths, dates16_deaths, dates17_deaths, dates18_deaths, dates19_deaths, dates20_deaths, dates21_deaths, dates22_deaths, dates23_deaths, dates24_deaths)

anxiety_value = indicators_US.narm$Value

#The compiled data has the date, the average number of cases for that date period, the average number of deaths for that date #period, and the reported CDC anxiety value for that data period. 
compiled_data = data.frame(dates, cases, deaths, anxiety_value)
compiled_data
```
```{r}
#Plot showing new cases vs anxiety value 
plot(x=cases, y=anxiety_value, main="cases ~ anxiety_value")  

#Plot showing new deaths vs anxiety value 
plot(x=deaths, y=anxiety_value, main="deaths ~ anxiety_value")
```

<span style='color:blue'>
In order to determine if there was any relationship between new COVID-19 cases and values for symptoms reported of Anxiety Disorder or Depressive Disorder, the relationship between new COVID-19 deaths and values for symptoms reported of Anxiety Disorder or Depressive Disorder were also examined. In roder to examine these relationships a linear model was fit. A linear model is appropriate for these analyses because the mdoel is explaining a continuous variable, $Y$, as a mathematical function of one or more variables, $X$, so that we can use this regression model to predict the $Y$ when only the $X$ is known.
The linear model is defined as $Y = \beta_1 + \beta_2X  + \epsilon$
where $Y$ represents the Anxiety Disorder or Depressive Disorder value. $\beta_1$ represents the intercept and $\beta_2$ represents the slope. $\epsilon$ represents the error term, the part of $Y$ that the regression model is unable to explain.
The $H_0$ is that $B_2=0$ and the $H_A$ is that $B_2\ne0$ meaning that there is a relationship between $X$ and $Y$. 
</span>

<span style='color:blue'>
This model assumes that the relationship between $X$ and $Y$ is linear and for ant fixed value of $X$, $Y$ is normally distributed. This model also assumes homoscedasticity, that the variance of residuals is the same for any value of $X$, and independence, that each observation is independent of each other. In order to put the values for $X$ and $Y$ on a similiar scale the log of $Y$ was used in the model. 
</span>

```{r}
linearMod_cases <- lm(log(cases) ~ anxiety_value, data=compiled_data)
summary(linearMod_cases)
```

<span style='color:blue'>
The reported summary for this model relating $Y$, number of new covid cases, to $X$ the value reported in the CDC Depression and Anxiety data, shows a positive linear relationship that is significant at p value = 0.05, from this information we can reject the null hypothesis, $H_0$, which states that $B_2=0$. 
</span>

```{r}
#Basic linear model relating new deaths vs anxiety value, the reported summary shows that the slope (depression_value) is not #significant and positive for this relationship. 
linearMod_deaths <- lm(log(deaths) ~ anxiety_value, data=compiled_data)
summary(linearMod_deaths)
```

<span style='color:blue'>
The reported summary for this model relating $Y$, number of new covid deaths, to $X$ the value reported in the CDC Depression and Anxiety data, shows a positive linear relationship that is not significant at p value = 0.05, from this information we can accept the null hypothesis, $H_0$, which states that $B_2=0$. 
</span>


# Sensitivity Analysis
Rubric:
- The report conducts extensive sentivity analysis to examine the plausibility of key assumptions in this analysis using visualization and rigorous testings.  In addition, results from alternative methods are obtained and compared with results in the main analysis,  to evaluate the impacts of key assumptions. 
- Results from the sensitivity analysis are interpreted in the context of the data set, and provide insights and suggestions for future analysis. Readers that are familiar with this topic or statistics gain confidence in the authors' analysis and knowledge. 


For our model we want to check the assumption of the linear regression model including:
* Linear Relationship
* Homoscedasticity
* Independence
* Normality 

```{r}
  ggplot(data = compiled_data, aes(x = anxiety_value, y=cases)) +
    geom_point(color = 'blue') +
    geom_smooth(method = "lm", se = FALSE, color = 'red')
```

The raw case data in relation to anxiety value does not appear to be linear. The lm() line calculation does not fit all of the data evenly and has some tailing at the high end of cases and anxiety values. While there does appear to be a correlation, the relationship does not appear to be linear. 

```{r}
  ggplot(data = compiled_data, aes(x = anxiety_value, y=log(cases))) +
    geom_point(color = 'blue') +
    geom_smooth(method = "lm", se = FALSE, color = 'red')
```

Upon an initial visualization to check for a linear relationship, we see that there is nothing to indicate otherwise. The data does not appear to be in any other form than linear. Log-transorming the data allows us to perform a linear regression as we cannot reject the assumption that the relationship is linear after tranforming the data.

```{r}
  plot(linearMod_cases)
```

In the residual plot, we want to check that the fitted residual line is mostly horizontal. We want our residual plot to have randomly distributed error terms with no clear pattern that are relatively close to this fitted line. In the model that relates log-transformed case counts and anxiety rates in the survey, we see this line is not perfectly horizontal, but nothing concerning to throw out the entire model. Some of the data is not spread evenly and there may be a couple outliers.

In the Normal Q-Q plot, we want to see that most of the data points are on the theoretical dashed line incicating normality. Our data here has a majority of the points near this line and has some tailing on either end of the distribution but nothing major. This would indicate normal distribution.

The Scale-Location plot, similar to the residual plot, wants a fitted line as horizontal as possible. Here we see similar results with no major deviations from a horizontal fitted line and a similar spread which should be looked into further via Cook's distance to determine outliers.

In the Residuals vs leverage plot with Cook's distance, we see that there are no points that pass the Cook's distance line. Since no points are outside the line our model assumptions hold.

### Insights and Suggestions for future Analysis

All our assumptions appear to hold for the model that related log-transformed case counts and anxiety rates in the country. While some data points come close to the Cook's distance line of 0.5 and our residual plot is not as great as it could be, there is not strong enough evidence in these plots to reject the model and it's assumptions. 

One way to be more conclusive about the assumptions is to have more data points, the data we collected is only weekly and the case-counts were aggregated weekly to allow for a proper comparison. More data points may be helpful, but weekly case data tends to be similar to daily case data and may not end up having that much of an impact on the conclusions. However, more data points would allow the model validation to be improved. 



# Conclusion
Rubric:
- The report clearly states the conclusions of the analysis in the context of the data set. The conlusion demonstrates thorough understanding of the model/method/procedure, real-world meaning of the findings, and acknowledgement of caveats of this report. 
- Readers that are familiar with this topic or statistics can understand the findings in the context of this topic, are aware of the caveats, and can make decisions based on the findings.

Our analysis finds that there is a relationship between COVID-Cases in the United States and anxiety rates. The model itself has a very low p-value which indicates that the null hypothesis of no relationship is rejected rather conclusively. In addition a high Adjusted $R^2$ of 0.8475 suggests the two observations are highly correlated. 

# References
Abbott, A. (2021). COVID’s mental-health toll: how scientists are tracking a surge in depression. Nature, 590(7845), 194–195. https://doi.org/10.1038/d41586-021-00175-z

Brodeur, A., Clark, A. E., Fleche, S., & Powdthavee, N. (2021). COVID-19, lockdowns and well-being: Evidence from Google Trends. Journal of Public Economics, 193, 104346. https://doi.org/10.1016/j.jpubeco.2020.104346

Bui, C. N., Peng, C., Mutchler, J. E., & Burr, J. A. (2020). Race and Ethnic Group Disparities in Emotional Distress Among Older Adults During the COVID-19 Pandemic. The Gerontologist, 61(2), 262–272. https://doi.org/10.1093/geront/gnaa217

Clemente-Suárez, V. J., Dalamitros, A. A., Beltran-Velasco, A. I., Mielgo-Ayuso, J., & Tornero-Aguilera, J. F. (2020). Social and Psychophysiological Consequences of the COVID-19 Pandemic: An Extensive Literature Review. Frontiers in Psychology, 11, 1–15. https://doi.org/10.3389/fpsyg.2020.580225

Delmastro, M., & Zamariola, G. (2020). Depressive symptoms in response to COVID-19 and lockdown: a cross-sectional study on the Italian population. Scientific Reports, 10(1), 1–10. https://doi.org/10.1038/s41598-020-79850-6

Fields JF, Hunter-Childs J, Tersine A, Sisson J, Parker E, Velkoff V, Logan C, and Shin H. Design and Operation of the 2020 Household Pulse Survey, 2020. U.S. Census Bureau. Forthcoming.
Ganson, K. T., Tsai, A. C., Weiser, S. D., Benabou, S. E., & Nagata, J. M. (2021). Job Insecurity and Symptoms of Anxiety and Depression Among U.S. Young Adults During COVID-19. Journal of Adolescent Health, 68(1), 53–56. https://doi.org/10.1016/j.jadohealth.2020.10.008

Holmes, E. A., O’Connor, R. C., Perry, V. H., Tracey, I., Wessely, S., Arseneault, L., Ballard, C., Christensen, H., Cohen Silver, R., Everall, I., Ford, T., John, A., Kabir, T., King, K., Madan, I., Michie, S., Przybylski, A. K., Shafran, R., Sweeney, A., … Bullmore, E. (2020). Multidisciplinary research priorities for the COVID-19 pandemic: a call for action for mental health science. The Lancet Psychiatry, 7(6), 547–560. https://doi.org/10.1016/s2215-0366(20)30168-1

